---
title: "Delta Speed"
author: "Anastasia Bernat"
date: "3/18/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
library(lme4)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

### Problems Noticing 

- qqplots of **all** models are not normal. Might need to retransform the distance_speed data either by standardizing it or performing another transformation. 


### Delta Speed

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R", # Loads and cleans data
                 "unique_flight_data.R",
                 "get_warnings.R", 
                 "compare_models.R",
                 "regression_output.R", # Cleans regression outputs; prints in color or B&W
                 "AICprobabilities.R",
                 "multinom_functions.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

output_col = FALSE # Change to TRUE if working in Base R or RStudio; FALSE if generating an HTML

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
d <- create_delta_data(data_tested)
```

#### Speed Diff ~ Mass Diff

```{r}
## Strong effect of mass
tidy_regression(lm(speed_diff ~ mass_diff, data=d), is_color = output_col) 
```


```{r}
hist(d$speed_diff) # much more normally distributed data and later will see the residuals pass the Shapiro-Wilk test
```

**No mass_diff**

```{r message=FALSE}
R = d$speed_diff
A = d$host_c
B = d$sex_c
C = d$sym_dist
D = d$mass_diff 
X = d$population 

data<-data.frame(R, A, B, C, D, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
model_comparisonsAIC(model_script)
```

```{r}
anova(m0, m2, test="Chisq") # Adding host matters
anova(m2, m6, test="Chisq") # Adding sym_dist does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ host_c + (1 | population), data=d, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

* no effect of host 

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-1, 0.1, s.test)
```

**Mass_diff**

```{r message=FALSE}
data <- data %>%
  filter(!is.na(D))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
model_comparisonsAIC(model_script)
```

```{r}
anova(m9, m14, test="Chisq") # Adding C does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ host_c + mass_diff + (1 | population), data=d, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

Strong negative effect of mass_diff where a gain in mass leads to slower speeds.

**wing2body**

```{r message=FALSE}
d <- d %>%
  filter(!is.na(mass_diff))

R = d$speed_diff
A = d$host_c
B = d$sex_c
C = d$wing2body_s
D = d$mass_diff 
X = d$population 

data<-data.frame(R, A, B, C, D, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
model_comparisonsAIC(model_script)
```
Same top model and wing2body makes no difference.


So what if it was only females, since they had the greatest mass changes?

### Delta Speed - Females Only

```{r}
d_fem <- d[d$sex == "F",] # low sample size 
```

**No mass_diff**

```{r message=FALSE}
R = d_fem$speed_diff
A = d_fem$host_c
B = d_fem$sym_dist
C = d_fem$mass_diff 
X = d_fem$population 

data<-data.frame(R, A, B, C, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 2-FF REMLF.R")
model_comparisonsAIC(model_script)
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ 1 + (1 | population), data=d_fem, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

**mass_diff**

```{r message=FALSE}
model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
model_comparisonsAIC(model_script)
```

```{r}
anova(m0, m1, test="Chisq") # strange. get error "refitting model(s) with ML (instead of REML)" even though these have been refit to REML=FALSE...
anova(m0, m3, test="Chisq") # Adding C does improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ mass_diff + (1 | population), data=d_fem, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

Strong negative effect of mass. 

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-1, 0.1, s.test)
```

**wing2body**

```{r message=FALSE}
R = d_fem$speed_diff
A = d_fem$host_c
B = d_fem$wing2body_s
C = d_fem$mass_diff 
X = d_fem$population 

data<-data.frame(R, A, B, C, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
model_comparisonsAIC(model_script)
```
```{r}
anova(m3, m5, test="Chisq") # Adding C marginally improves fit
anova(m5, m9, test="Chisq") # Adding A*C does not improve fit
anova(m3, m6, test="Chisq") # Adding B does not improve fit
```
```{r}
speed_model <- lmer(speed_diff ~ host_c + mass_diff + (1 | population), data=d_fem, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

Strong negative effect of mass. 
Marginal negative effect of host.

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-1, 0.1, s.test)
```

**eggs**

```{r}
R = d_fem$speed_diff
A = d_fem$host_c
B = d_fem$egg_case
C = d_fem$mass_diff 
X = d_fem$population 

data<-data.frame(R, A, B, C, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
model_comparisonsAIC(model_script)
```

```{r}
anova(m5, m7, test="Chisq") # Adding B does not improve fit
anova(m5, m9, test="Chisq")
```
Egg case does not matter.

### Plots

```{r}
plot(d$mass_diff, d$speed_diff, 
     xlab="Changes in Mass between T1 and T2 (g)",
     ylab="Changes in Speed between T1 and T2 (m/s)")
```




