---
title: 'Modeling Flight Maximum Speed'
author: "Anastasia Bernat"
date: "12/16/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Winter_2020/stats/")

library(lme4)
library(rethinking)

library(dplyr)
library(tidyselect)
library(stringr)

library(glmnet)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(randomcoloR)
library(rstatix)
library(ggpubr)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials: Speed Flight Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (glm) and mixed effect modeling (glmer) to analyze the flight results.

### Cleaning the Data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R" # Loads and cleans data
                  )

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_flew_all <- data_tested[data_tested$flew_b == 1, ] # remove who didn't fly
```

```{r}
data_flew_all = read.csv("data/flight_stats_summary5.csv")
data_flew_all <- data_flew_all[data_flew_all$average_speed != 0, ]

par(mfrow=c(2,2))
hist(data_flew_all$max_speed1,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
hist(data_flew_all$max_speed2,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
hist(data_flew_all$max_speed3,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
hist(data_flew_all$max_speed4,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
hist(data_flew_all$max_speed10,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
```

```{r fig.height=10, fig.width=10}
small = data_flew_all %>%
  select(max_speed1, max_speed2, max_speed3, max_speed4)


v = c(1,2,3,4)
plot(v, unlist(c(data_flew_all[1,15:18])), ylim=c(0,0.8), type="l")
for (i in seq(2,93, 1)) {
  points(v, unlist(c(data_flew_all[i,15:18])), type="l")
}
```


```{r}
par(mfrow=c(1,2))
hist(data_flew_all$max_speed1,   # not normally distributed
     main="Max Speed Histogram", 
     xlab="Max Speed (m/s)", 
     ylim=c(0,100))
# boxplot(max_speed1~sex,data=data_flew_all, 
#         main="Max Speed by Sex Boxplot",
#         ylab="Max Speed (m/s)", 
#         xlab="Sex")
```

```{r}
sort(data_flew_all$max_speed1) # need to do flight trajectories to see *where* these pop up
```

```{r}
# pwc = "Pairwise Wilcoxon test" between groups
pwc <- data_flew_all %>% 
  dunn_test(max_speed1 ~ sex, p.adjust.method = "bonferroni") %>%
  add_xy_position(x = "sex")
res.kruskal <- data_flew_all %>% kruskal_test(max_speed1 ~ sex)

ggplot(data_flew_all, aes(sex, max_speed1)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE), 
    caption = get_pwc_label(pwc)
    ) + 
  theme(legend.position="none") +
  xlab("Sex") +
  ylab("Max speed (m/s)") +
  geom_boxplot(width=0.1) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```
There's a difference in max speeds between the sexes where females can reach faster max speeds. 
